---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




## Examples

- choropleth
- isochrones
- polygons, polylines, shapes, 
- drawing / editing
- 2d vs 2.5d

- static & interactive

- shiny integration

- googleway (requires API key)
- mapdeck   (requires API key)
- (leaflet)
- (ggmap)
- (mapview)

## googleway

```{r visualisationGoogleway} 
library(googleway)
set_key(read.dcf("~/Documents/.googleAPI", fields = "GOOGLE_MAP_KEY"))
```

The `googleway` library provides access to Google Maps API, both for creating maps, and using the mapping services such as geocoding and searching for places. 

---

Given it uses Google Maps you can also use their streetview serivce


## Mapdeck

```{r visualisationMapdeck}
library(mapdeck)
set_token(read.dcf("~/Documents/.googleAPI", fields = "MAPBOX"))
```

The `mapdeck` library provides access to Uber's `Deck.gl` framework, and uses `Mapbox GL` as the mapping layer. Deck.gl's use of WebGL enables it to render large datasets (up to millions of points). It also provides a '2.5d' perspective of data, where the user can tilt, drag and rotate the map (as well as pan and zoom), and view data 'extruded' from the map

```{r mapdeckExtrudedExample}

```

## Shiny

All the interactive maps listed are built using the [`htmlwidgets`](https://www.htmlwidgets.org/) framework. Which is natively supported inside [`Shiny`](https://shiny.rstudio.com/) applications. This means you can include the maps in an interactive dashboard.

```{r visualisationShinyExample}


```