---
title: "Choropleth - stroke by postcode"
author: "Richard Beare"
date: "08/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::read_chunk("mmc_surrounds.R")
```

# Introduction

Start of the choropleth example. Choropleths are a staple of geospatial visualization. The idea of this 
example is to introduced those basics in stroke context. I've sketched a working outline with simple
features, census data and incidence data from a paper.

The example loads the data, estimates annual stroke incidence, performs a join, calculates some
geospatial measures, filters postcodes based on distance to the hospital and displays the results
with tmap.

Display can be improved (a pin for this hospital), and other packages can be used in addition to display alternatives.


```{r LibrariesAndCensusData, message=FALSE, warning=FALSE}

```

Could geocode this
```{r MonashMedicalCentre}

```


```{r StrokeIncidence}

```

Combining demographics and spatial data

```{r JoinCensusAndBoundaries}

```

```{r SpatialComputations}

```

```{r FilteringPostcodes}

```

```{r InteractiveDisplay}

```

