---
title: "Catchment areas for rehabilitation centres"
output:
  rmarkdown::html_vignette:
    self_contained: no

  md_document:
    variant: markdown_github
---

```{r opts, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = TRUE,
  message = TRUE,
  width = 120,
  comment = "#>",
  fig.retina = 2,
  fig.path = "README-"
)
```

## ---- Libraries and Census Data

```{r initial-load}
library(tidyverse)
library(sf)
library(units)
library(tmaptools)
postcodeboundariesAUS <- 
    file.path(here::here(), "ABSData", "Boundaries/POA_2016_AUST.shp") %>%
    sf::read_sf ()

basicDemographicsVIC <- file.path(here::here(), "ABSData",
                                  "2016 Census GCP Postal Areas for VIC",
                                  "2016Census_G01_VIC_POA.csv") %>%
    readr::read_csv()
```

Clean up the demographics to only those columns that we're interested in.
Presume just for illustrative purposes here that those are only the basic "Age"
classes. There are also columns about the ages of persons attending educational
institutions which need to be removed.
```{r initial-clean}
library(magrittr)

basicDemographicsVIC <- select(basicDemographicsVIC, POA_CODE_2016, starts_with("Age_"), -starts_with("Age_psns_"))
    
```



## ---- JoinCensusAndBoundaries ----

Join the demographics and shape tables, retaining victoria only use postcode
boundaries as the reference data frame so that coordinate reference system is
retained.
```{r join-demo-postcodes}
basicDemographicsVIC <- right_join(postcodeboundariesAUS,
                                   basicDemographicsVIC, 
                                   by=c("POA_CODE" = "POA_CODE_2016"))
```

## ---- GeocodeRehabNetwork ----

To be clean up
```{r rehab-addresses}
rehab_addresses <- c(DandenongHospital = "Dandenong Hospital, Dandenong VIC 3175, Australia",
                     CaseyHospital = "62-70 Kangan Dr, Berwick VIC 3806, Australia",
                     KingstonHospital = "The Kingston Centre, Heatherton VIC 3202, Australia")
RehabLocations <- tmaptools::geocode_OSM(rehab_addresses, as.sf=TRUE)
```

transform rehab locations to the same reference system
```{r rehab-addresses-transform}
RehabLocations <- sf::st_transform(RehabLocations,
                                   sf::st_crs(basicDemographicsVIC))
```

## Check geocoding

With `tmap`:
```{r tmap1, eval = FALSE}
library(tmap)
tmap_mode("view")

tm_shape(RehabLocations) + tm_markers() + 
  tm_basemap("OpenStreetMap")
```
Or with `mapdeck`:
```{r mapdeck1, eval = FALSE}
library(mapdeck)
set_token(Sys.getenv("MAPBOX_TOKEN"))
mapdeck(location = c(145.2, -38),
        zoom = 12) %>%
    add_pointcloud (RehabLocations,
        layer_id = "rehab-locations")
```

## ---- Postcodes surrounding rehab locations

There are 699 postcodes which we now want to reduce to only those within a
specified distance of the rehab locations, chosen here as 10km. Note that we
just use straight line distances here, because we only need to roughly determine
which postcodes surround our rehab centres. The subsequent calculations will
then use more accurate distances along street networks.

```{r postcode-dists}
dist_to_loc <- function (geometry, location){
    units::set_units(st_distance(geometry, location)[,1], km)
}
dist_range <- units::set_units(10, km)

#basicDemographicsVIC <- basicDemographicsVIC_old
basicDemographicsVIC_old <- basicDemographicsVIC
basicDemographicsVIC <- mutate(basicDemographicsVIC,
       DirectDistanceToDandenong = dist_to_loc(geometry,RehabLocations["DandenongHospital", ]),
       DirectDistanceToCasey     = dist_to_loc(geometry,RehabLocations["CaseyHospital", ]),
       DirectDistanceToKingston  = dist_to_loc(geometry,RehabLocations["KingstonHospital", ]),
       DirectDistanceToNearest   = pmin(DirectDistanceToDandenong,
                                        DirectDistanceToCasey,
                                        DirectDistanceToKingston)
)
basicDemographicsRehab <- filter(basicDemographicsVIC,
                                 DirectDistanceToNearest < dist_range) %>%
        mutate(Postcode = as.numeric(POA_CODE16)) %>%
        select(-starts_with("POA_"))
```
That reduces the data down to 47 nearby postcodes, with the last 2 lines
converting all prior postcode columns (of which there were several all beginning
with "POA") to a single numeric column named "Postcode".


## ---- SamplePostCodes ----
Select random addresses using a geocoded database

```{r PSMA-install, eval = FALSE}
devtools::install_github("HughParsonage/PSMA")
```

Will increase this for the real example
```{r addressesPerPostcode}
addressesPerPostcode <- 500
```

A special function so we can sample the postcodes as we go.  Sampling syntax is
due to the use of data.table inside PSMA. The last `st_as_sf()` command converts
the points labelled "LONGITUDE" and "LATITUDE" into `sf::POINT` objects. (This
function takes a few seconds because of the `fetch_postcodes` call.)
```{r random-addresses}
library(PSMA)
samplePCode <- function(pcode, number) {
  d <- fetch_postcodes(pcode)
  return(d[, .SD[sample(.N, min(number, .N))], by=.(POSTCODE)])
}

randomaddresses <- map(basicDemographicsRehab$Postcode,
                       samplePCode,
                       number=addressesPerPostcode) %>%
            bind_rows() %>%
            sf::st_as_sf(coords = c("LONGITUDE", "LATITUDE"),
                         crs=st_crs(basicDemographicsRehab),
                         agr = "constant")
```


## ---- PlotSampleLocations ----

With `tmap`:
```{r tmap2, eval = FALSE}
library(tmap)
tmap_mode("view")
tm_shape(randomaddresses) + tm_markers(clustering=FALSE) + 
    tm_basemap("OpenStreetMap")
head(randomaddresses)
```
or with `mapdeck`
```{r mapdeck2, eval = FALSE}
library(mapdeck)
set_token(Sys.getenv("MAPBOX_TOKEN"))
mapdeck(location = c(145.2, -38),
        zoom = 12) %>%
    add_pointcloud (randomaddresses,
                    radius = 2,
                    layer_id = "randomaddresses")
```


## ---- AddressesToRehab ----

Compute the road distance and travel time from each address to each hospital.
This first requires a local copy of the street network within the bounding
polygon defined by `basicDemographicsRehab`. This is easiest done with the
`dodgr` package, which directly calls the `osmdata` package to do the
downloading.

### Street Network

It is instructive to examine the `mapdeck` view of the postcode polygons:
```{r mapdeck-postcode-polygons, eval = FALSE}
library(mapdeck)
set_token(Sys.getenv("MAPBOX_TOKEN"))
mapdeck(location = c(145.2, -38),
        zoom = 12) %>%
    add_polygon (basicDemographicsRehab, fill_colour="Postcode",
        layer_id = "randomaddresses")
```
The basic way to download the street network is within a defined, implicitly
rectangular, bounding box, but in this case that extends from Mornington to St
Kilda, and out to the Dandenongs, and even Koo Wee Rup! It is much better to
extract the street network only within the polygon defining our nearby postcode
areas, which first needs to be re-projected onto the CRS of OpenStreetMap data,
which is epsg4326. `st_union` merges all of the polygons to form the single
enclosing polygon, and the final command simply extracts the longitudinal and
latitudinal coordinates of that polygon (rather than leaving them in `sf`
format).
```{r postcode-bounding-polygon, eval = TRUE}
bounding_polygon <- sf::st_transform(basicDemographicsRehab,
                                     sf::st_crs(4326)) %>%
    sf::st_union () %>%
    sf::st_coordinates ()
bounding_polygon <- bounding_polygon [, 1:2]
```
We can now download the street network enclosed within that polygon. Note that
this is still a rather large network - over 40MB of data representing over
60,000 street sections - that might take a minute or two to process. It is
therefore easier to save the result to disc for quicker re-usage.

```{r get-streetnet, eval = FALSE}
library(dodgr)
system.time (
dandenong_streets <- dodgr_streetnet (bounding_polygon, expand = 0, quiet = FALSE)
)
saveRDS (dandenong_streets, file = "dandenong-streets.Rds")
format (file.size ("dandenong-streets.Rds"), big.mark = ",")
nrow (dandenong_streets)
```
The network can then be re-loaded with
```{r reload-streetnet-demo, eval = FALSE}
dandenong_streets <- readRDS ("dandenong-streets.Rds")
```

### Distances to Hospitals

The `dodgr` package needs to de-compose the `sf`-formatted street network, which
consists of long, connected road segments, into individual edges. This is done
with the `weight_streetnet()` function, which modifies the distance of each edge
to reflect typical travel conditions for a nominated mode of transport.

```{r load-streetnet}
library (dodgr)
dandenong_streets <- readRDS ("dandenong-streets.Rds")
net <- weight_streetnet (dandenong_streets, wt_profile = "motorcar")
nrow (dandenong_streets); nrow (net)
```
The resultant network has a `d_weighted` column which preferentially weights the
distances for the nominated mode of tranport. Those parts of the network which
are unsuitable for vehicular transport have values of `.Machine$double.xmax = `
`r .Machine$double.xmax`. Because we want to align our random points to the
*routable* component of the network, these need to be removed.
```{r}
net <- net [which (net$d_weighted < .Machine$double.xmax), ]
nrow (net)
```

The `r nrow(dandenong_streets)` streets have been converted to `r nrow(net)`
distinct edges. We can now use the `net` object to calculate the distances,
along with simple numeric coordinates of our routing points, projected on to the
same CRS as OpenStreetMap (OSM), which is 4326:
```{r get-routing-coordinates}
fromCoords <- st_coordinates (st_transform (randomaddresses, crs = 4326))
toCoords <- st_coordinates (st_transform (RehabLocations, crs = 4326))
```
Although not necessary, distance calculation is quicker if we map these `from`
and `to` points precisely on to the network itself. OSM assigns unique
identifiers to every single object, and so our routing coordinates can be
converted to OSM identifiers of the nearest street nodes. The nodes themselves
are obtained with the `dodgr_vertices()` function.
```{r convert-routing-coordinates}
nodes <- dodgr_vertices (net)
fromIDX <- match_pts_to_graph (nodes, fromCoords, connected = TRUE)
from <- unique (nodes$id [fromIDX])
to <- nodes$id [match_pts_to_graph (nodes, toCoords, connected = TRUE)]
```
The matrices of `from` and `to` coordinates have now been converted to simple
vectors of OSM identifiers.

```{r dodgr-dists}
system.time (
             d <- dodgr_dists (net, from = from, to = to)
)
```
And that takes only around 1 second to calulate distances between (3 rehab
centres times 20,000 random addresses = ) 60,000 pairs of points.



## ---- CatchmentBasins ----

First assign each point to its nearest hospital according to the street network
distances returned from `dodgr_dists`. Note that points on the outer periphery
of the network may not necessarily be connected, as we'll see below.
```{r alllocate-points}
DestNames <- c(rownames(RehabLocations), "Disconnected")
# assign each source address to the nearest destination
DestNumber <- as.numeric (apply(d, MARGIN=1, which.min))
DestNumber [is.na (DestNumber)] <- 4 # the disconnected points
BestDestination <- DestNames[DestNumber]
table (BestDestination)
```
And there are `r length (which (DestNumber == 4))` points that are not
connected. The allocation of points, including these disconnected ones, can be
inspected on a map with the following code, start by setting up a `data.frame`
of `fromCoords`.
```{r fromCoords}
fromCoords <- nodes [match (from, nodes$id), ]
fromCoords$DestNumber <- DestNumber
fromCoords$Destination <- BestDestination
```
```{r, eval = FALSE}
library(mapdeck)
set_token(Sys.getenv("MAPBOX_TOKEN"))
mapdeck(location = c(145.2, -38),
        zoom = 10) %>%
    add_pointcloud (data = fromCoords, lon = "x", lat = "y",
                    fill_colour = "DestNumber",
                    palette = "plasma")
```
As a final step, we'll convert those clusters of points into enclosing polygons,
using a Voronoi tesselation. `sf::st_voronoi` doesn't return the polygons in the
same order as the original points, requiring a manual re-sorting in order to use
this to match voronoi polygons to points for each catchment.
```{r voronoi}
g <- st_multipoint(as.matrix(fromCoords[,c("x", "y")]))
v <- st_voronoi(x=g) # results in geometry collection objects
v <- st_collection_extract(v) # converts to polygons
fromCoords_sf <- st_as_sf(fromCoords, coords=c("x", "y"))
vorder <- unlist(st_intersects(fromCoords_sf, v))
v <- v[vorder] # polygons in same order as points
v <- st_sf (DestNumber = fromCoords$DestNumber,
            Destination = fromCoords$Destination,
            geometry = v,
            crs = 4326)
```
We then just need to filter those Voronoi polygons associated with each
catchment, and extract the surrounding polygons. (This can take quite some
time.)
```{r catchment-polygons, eval = FALSE}
bounding_polygon <- sf::st_transform(basicDemographicsRehab,
                                     sf::st_crs(4326)) %>%
  sf::st_union () 
v <- lapply (1:3, function (i) {
                 v [v$DestNumber == i, ] %>%
                     st_intersection (bounding_polygon) %>%
                     st_union() })
v <- st_sf (DestNumber = 1:3,
            Destination = DestNames [1:3],
            geometry = do.call (c, v))
```

Then plot with `mapdeck`:
```{r catchment-polygon-plot, eval = FALSE}
mapdeck(location = c(145.2, -38),
        zoom = 10) %>%
    add_polygon (data = v,
                    fill_colour = "DestNumber",
                    fill_opacity = 150,
                    palette = "plasma") %>%
    add_pointcloud (data = RehabLocations)
```


## ---- CasesPerCentre ----

Also need a per postcode breakdown of proportion of addresses going to each
centre, so that we can compute the number of cases going to each centre. The
above procedure occasionally mapped multiple addresses onto the same network
points. As we were only interested in the enclosing polygons, these repeated
points were removed. In the present case, however, these repeats need to be
counted, so we need to go back to where we were before, which is the
`randomaddresses`.

```{r}
dim (randomaddresses); dim (fromCoords)
length (from); length (DestNumber)
```
We need to repeat the calculation of `DestNumber` using the full set of
`fromCoords`, including those repeatedly matched onto same network points.
```{r redo-destinations}
fromCoords <- st_coordinates (st_transform (randomaddresses, crs = 4326))
fromIDX <- match_pts_to_graph (nodes, fromCoords, connected = TRUE)
from <- nodes$id [fromIDX]
to <- nodes$id [match_pts_to_graph (nodes, toCoords, connected = TRUE)]
d <- dodgr_dists (net, from = from, to = to)
DestNames <- c(rownames(RehabLocations), "Disconnected")
DestNumber <- as.numeric (apply(d, MARGIN=1, which.min))
DestNumber [is.na (DestNumber)] <- 4 # the disconnected points
BestDestination <- DestNames[DestNumber]
```
The following lines then group the above data by both postcode and destination
hospital.
```{r postcode-groups}
postcodes <- data.frame (POSTCODE = randomaddresses$POSTCODE,
                         DestNumber = DestNumber,
                         Destination = BestDestination,
                         stringsAsFactors = FALSE) %>%
    group_by (POSTCODE, DestNumber, Destination) %>%
    summarise (n = length (DestNumber))
postcodes
```

