---
title: "Catchment areas for rehabilitation centres"
output:
  rmarkdown::html_vignette:
    self_contained: no

  md_document:
    variant: markdown_github
---

```{r opts, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = TRUE,
  message = TRUE,
  width = 120,
  comment = "#>",
  fig.retina = 2,
  fig.path = "README-"
)
```

## ---- Libraries and Census Data

```{r initial-load}
library(tidyverse)
library(sf)
library(units)
library(tmaptools)
postcodeboundariesAUS <- 
    file.path(here::here(), "ABSData", "Boundaries/POA_2016_AUST.shp") %>%
    sf::read_sf ()

basicDemographicsVIC <- file.path(here::here(), "ABSData",
                                  "2016 Census GCP Postal Areas for VIC",
                                  "2016Census_G01_VIC_POA.csv") %>%
    readr::read_csv()
```

Clean up the demographics to only those columns that we're interested in.
Presume just for illustrative purposes here that those are only the basic "Age"
classes. There are also columns about the ages of persons attending educational
institutions which need to be removed.
```{r initial-clean}
library(magrittr)
nms <- names (basicDemographicsVIC)
keep_cols <- c ("POA_CODE_2016", "Age")
remove_cols <- c ("Age_psns")
keep_index <- sapply (keep_cols, function (i) grep (i, nms)) %>%
    unlist() %>%
    unique()
remove_index <- sapply (remove_cols, function (i) grep (i, nms)) %>%
    unlist () %>%
    unique ()
keep_index <- keep_index [!keep_index %in% remove_index]
basicDemographicsVIC %<>%
    select(keep_index) %>%
    select(-remove_index)
```



## ---- JoinCensusAndBoundaries ----

Join the demographics and shape tables, retaining victoria only use postcode
boundaries as the reference data frame so that coordinate reference system is
retained.
```{r join-demo-postcodes}
basicDemographicsVIC <- right_join(postcodeboundariesAUS,
                                   basicDemographicsVIC, 
                                   by=c("POA_CODE" = "POA_CODE_2016"))
```

## ---- GeocodeRehabNetwork ----

To be clean up
```{r rehab-addresses}
rehab_addresses <- c(DandenongHospital = "Dandenong Hospital, Dandenong VIC 3175, Australia",
                     CaseyHospital = "62-70 Kangan Dr, Berwick VIC 3806, Australia",
                     KingstonHospital = "The Kingston Centre, Heatherton VIC 3202, Australia")
RehabLocations <- tmaptools::geocode_OSM(rehab_addresses, as.sf=TRUE)
```

transform rehab locations to the same reference system
```{r rehab-addresses-transform}
RehabLocations <- sf::st_transform(RehabLocations,
                                   sf::st_crs(basicDemographicsVIC))
```

## Check geocoding

With `tmap`:
```{r tmap1, eval = FALSE}
library(tmap)
tmap_mode("view")

tm_shape(RehabLocations) + tm_markers() + 
  tm_basemap("OpenStreetMap")
```
Or with `mapdeck`:
```{r mapdeck1, eval = FALSE}
library(mapdeck)
set_token(Sys.getenv("MAPBOX_TOKEN"))
mapdeck(location = c(145.2, -38),
        zoom = 12) %>%
    add_pointcloud (RehabLocations,
        layer_id = "rehab-locations")
```

## ---- Postcodes surrounding rehab locations

There are 699 postcodes which we now want to reduce to only those within a
specified distance of the rehab locations, chosen here as 10km. Note that we
just use straight line distances here, because we only need to roughly determine
which postcodes surround our rehab centres. The subsequent calculations will
then use more accurate distances along street networks.

```{r postcode-dists}
dist_to_loc <- function (geometry, location){
    units::set_units(st_distance(geometry, location)[,1], km)
}
dist_range <- units::set_units(10, km)

#basicDemographicsVIC <- basicDemographicsVIC_old
basicDemographicsVIC_old <- basicDemographicsVIC
basicDemographicsVIC <- mutate(basicDemographicsVIC,
       DirectDistanceToDandenong = dist_to_loc(geometry,RehabLocations["DandenongHospital", ]),
       DirectDistanceToCasey     = dist_to_loc(geometry,RehabLocations["CaseyHospital", ]),
       DirectDistanceToKingston  = dist_to_loc(geometry,RehabLocations["KingstonHospital", ]),
       DirectDistanceToNearest   = pmin(DirectDistanceToDandenong,
                                        DirectDistanceToCasey,
                                        DirectDistanceToKingston)
)
basicDemographicsRehab <- filter(basicDemographicsVIC,
                                 DirectDistanceToNearest < dist_range) %>%
        mutate(Postcode = as.numeric(POA_CODE16)) %>%
        select(-grep("POA_", names(basicDemographicsVIC)))
```
That reduces the data down to 47 nearby postcodes, with the last 2 lines
converting all prior postcode columns (of which there were several all beginning
with "POA") to a single numeric column named "Postcode".


## ---- SamplePostCodes ----
Select random addresses using a geocoded database

```{r PSMA-install, eval = FALSE}
devtools::install_github("HughParsonage/PSMA")
```

Will increase this for the real example
```{r addressesPerPostcode}
addressesPerPostcode <- 500
```

A special function so we can sample the postcodes as we go.  Sampling syntax is
due to the use of data.table inside PSMA. The last `st_as_sf()` command converts
the points labelled "LONGITUDE" and "LATITUDE" into `sf::POINT` objects. (This
function takes a few seconds because of the `fetch_postcodes` call.)
```{r random-addresses}
library(PSMA)
samplePCode <- function(pcode, number) {
  d <- fetch_postcodes(pcode)
  return(d[, .SD[sample(.N, min(number, .N))], by=.(POSTCODE)])
}

randomaddresses <- map(basicDemographicsRehab$Postcode,
                       samplePCode,
                       number=addressesPerPostcode) %>%
            bind_rows() %>%
            sf::st_as_sf(coords = c("LONGITUDE", "LATITUDE"),
                         crs=st_crs(basicDemographicsRehab),
                         agr = "constant")
```


## ---- PlotSampleLocations ----

With `tmap`:
```{r tmap2, eval = FALSE}
library(tmap)
tmap_mode("view")
tm_shape(randomaddresses) + tm_markers(clustering=FALSE) + 
    tm_basemap("OpenStreetMap")
head(randomaddresses)
```
or with `mapdeck`
```{r mapdeck2, eval = FALSE}
library(mapdeck)
set_token(Sys.getenv("MAPBOX_TOKEN"))
mapdeck(location = c(145.2, -38),
        zoom = 12) %>%
    add_pointcloud (randomaddresses,
                    radius = 2,
                    layer_id = "randomaddresses")
```


## ---- AddressesToRehab ----

Compute the road distance and travel time from each address to each hospital.
This first requires a local copy of the street network within the bounding
polygon defined by `basicDemographicsRehab`. This is easiest done with the
`dodgr` package, which directly calls the `osmdata` package to do the
downloading.

### Street Network

It is instructive to examine the `mapdeck` view of the postcode polygons:
```{r mapdeck-postcode-polygons, eval = FALSE}
library(mapdeck)
set_token(Sys.getenv("MAPBOX_TOKEN"))
mapdeck(location = c(145.2, -38),
        zoom = 12) %>%
    add_polygon (basicDemographicsRehab,
        layer_id = "randomaddresses")
```
The basic way to download the street network is within a defined, implicitly
rectangular, bounding box, but in this case that extends from Mornington to St
Kilda, and out to the Dandenongs, and even Koo Wee Rup! It is much better to
extract the street network only within the polygon defining our nearby postcode
areas, which first needs to be re-projected onto the CRS of OpenStreetMap data,
which is epsg4326. `st_union` merges all of the polygons to form the single
enclosing polygon, and the final command simply extracts the longitudinal and
latitudinal coordinates of that polygon (rather than leaving them in `sf`
format).
```{r postcode-bounding-polygon, eval = FALSE}
bounding_polygon <- sf::st_transform(basicDemographicsRehab,
                                     sf::st_crs(4326)) %>%
    sf::st_union () %>%
    sf::st_coordinates ()
bounding_polygon <- bounding_polygon [, 1:2]
```
We can now download the street network enclosed within that polygon. Note that
this is still a rather large network - over 40MB of data representing over
60,000 street sections - that might take a minute or two to process. It is
therefore easier to save the result to discuss for quicker re-usage.

```{r get-streetnet, eval = FALSE}
library(dodgr)
system.time (
dandenong_streets <- dodgr_streetnet (bounding_polygon, expand = 0, quiet = FALSE)
)
saveRDS (dandenong_streets, file = "dandenong-streets.Rds")
format (file.size ("dandenong-streets.Rds"), big.mark = ",")
nrow (dandenong_streets)
```
The network can then be re-loaded with
```{r reload-streetnet-demo, eval = FALSE}
dandenong_streets <- readRDS ("dandenong-streets.Rds")
```

### Distances to Hospitals

The `dodgr` package needs to de-compose the `sf`-formatted street network, which
consists of long, connected road segments, into individual edges. This is done
with the `weight_streetnet()` function, which modifies the distance of each edge
to reflect typical travel conditions for a nominated mode of transport.

```{r load-streetnet}
library (dodgr)
dandenong_streets <- readRDS ("dandenong-streets.Rds")
net <- weight_streetnet (dandenong_streets, wt_profile = "motorcar")
nrow (dandenong_streets); nrow (net)
```
The 62,624 streets have been converted to 626,391 distinct edges. We can now use
the `net` object to calculate the distances, along with simple numeric
coordinates of our routing points, projected on to the same CRS as
OpenStreetMap (OSM), which is 4326:
```{r get-routing-coordinates}
from <- st_coordinates (st_transform (randomaddresses, crs = 4326))
to <- st_coordinates (st_transform (RehabLocations, crs = 4326))
```
Although not necessary, distance calculation is quicker if we map these `from`
and `to` points precisely on to the network itself. OSM assigns unique
identifiers to every single object, and so our routing coordinates can be
converted to OSM identifiers of the nearest street nodes. The nodes themselves
are obtained with the `dodgr_vertices()` function.
```{r convert-routing-coordinates}
nodes <- dodgr_vertices (net)
from <- nodes$id [match_pts_to_graph (nodes, from, connected = TRUE)]
to <- nodes$id [match_pts_to_graph (nodes, to, connected = TRUE)]
```
The matrices of `from` and `to` coordinates have now been converted to simple
vectors of OSM identifiers.

```{r dodgr-dists}
system.time (
             d <- dodgr_dists (net, from = from, to = to)
)
```
And that takes only around 1 second to calulate distances between (3 rehab
centres times 20,000 random addresses = ) 60,000 pairs of points.



## ---- CatchmentBasins ----

I did this by creating a Voronoi tesselation, of the random addresses, and
combining the regions according to the nearest hospital. Hopefully there are sf
tools for this.

## ---- CasesPerCentre ----

Also need a per postcode breakdown of proportion of addresses going to each
centre, so that we can compute the number of cases going to each centre
